---
title: "Homework 1 (p8105)"
author: "Sanyu Takirambudde (stt2133)"
date: 2026-09-26
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Problem 1

```{r}
data("penguins", package = "palmerpenguins")
```

### Penguins Dataset Description

**Variables:** The variables included in the penguins dataset are *species, island, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex, and year*. 

**Size:** The dataset contains `r nrow(penguins)` rows and `r ncol(penguins)` columns. 

**Mean Flipper Length:** The mean flipper length across all penguins in the dataset is `r round(mean(pull(drop_na(penguins, flipper_length_mm), flipper_length_mm)), 2)` mm.

### Scatterplot of Flipper Length (mm) vs Bill Length (mm)

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + 
  geom_point()
ggsave("p8105_hw1_prob1_scatterplot.png")
```

# Problem 2

```{r}
set.seed(1)
hw1_prob2_df = 
  tibble(
    std_norm_samp = rnorm(10, mean = 0, sd = 1),
    std_norm_samp_pos = std_norm_samp > 0,
    vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    vec_factor = factor(c("low", "medium", "high", "low", "medium", "high", "low", "medium", "high", "low"))
  )
```

```{r}
mean(pull(hw1_prob2_df, std_norm_samp))
mean(pull(hw1_prob2_df, std_norm_samp_pos))
mean(pull(hw1_prob2_df, vec_char))
mean(pull(hw1_prob2_df, vec_factor))
```

**Variables that work in taking the mean:** The random sample a standard Normal distribution (std_norm_samp) and the logical vector indicating whether elements of the sample are greater than 0 (std_norm_samp_pos).

**Variables that do not work in taking the mean:** The character vector (vec_char) and factor vector with 3 different factor “levels" (vec_factor).
